
---
title: "Daily Classroom Occupancy"
author: "Who's There"
header-includes: 
   \usepackage{graphicx}
   \usepackage{fancyhdr}
   \pagestyle{fancy}
   \setlength\headheight{50pt}
   \fancyhead[L]{\includegraphics[width=2cm]{./logo.png} Who's There}
output:
  pdf_document:
    fig_height: 6
    fig_width: 16
---

```{r, echo=FALSE,results='hide', message=FALSE, warning=FALSE}
#echo=FALSE,results='hide', message=FALSE, warning=FALSE
library(httr)
library(jsonlite)
library(lubridate)
library(ggplot2) #package for making graphs
source("http://peterhaschke.com/Code/multiplot.R") #for using multiplot
options(stringsAsFactors = FALSE)#turn off the auto transformation of string in factors
library(dplyr) 
library(tidyjson)
library(magrittr)
library(data.table)
library(digest)
library(Rcpp)


#set up the url and the path of the api for server
#url <- "paste("http://csi420-01-vm1.ucd.ie:8080"/api/data?request=Room_no&Room_no=B003&request2=Date&Date=2015-11-03"

#get the variable from command line
args <- commandArgs(trailingOnly = TRUE)

#when run the R script you have to specify the name of the file and the arguments
#in windows will be: Rscript --vanilla namescript.R arg1 arg2
#assign a name to the 2 passed argument
room <- args[1]
date_input <- args[2]

#set up the url and the path of the api locally
url <- "http://csi420-01-vm1.ucd.ie:8080/api/data?request=Room_no&Room_no=%s&request2=Date&Date=%s"

#set up the url and the path of the api locally
url <- sprintf(url, room, date_input)
# for test only: print(paste("http://192.168.10.2:8080/api/data?request=Room_no&Room_no=",room,"&request2=Date&Date=",date)) 

raw_contents <- GET(url = url)

json_raw <- httr::content(raw_contents, type = "text", encoding = "UTF-8")

tidy_json <- json_raw %>% as.tbl_json

#Getting the detail for printing the main information on the PDF 
details <-
  tidy_json %>%
  enter_object("Room_no") %>%
  enter_object(room) %>%
  spread_values(Campus = jstring("Campus"),Building = jstring("Building"), Capacity=jnumber("Capacity"), Plug_friendly= jstring("Plug_friendly"), Floor =jstring("Floor_no")) 

#Getting the keys for Timeslot
time <- 
  tidy_json %>%
  enter_object("Room_no") %>%
  enter_object(room) %>%
  enter_object("Date") %>%
  enter_object(date_input) %>%
  enter_object("Timeslot") %>%
  gather_keys 

#for loop for gathering the information regarding the estimates on room occupancy
#made by the model in each time slot and storing them in a dataframe 
  for (i in 1:8){
  occupancy<-
  tidy_json %>%
  enter_object("Room_no") %>%
  enter_object(room) %>%
  enter_object("Date") %>%
  enter_object(date_input) %>%
  enter_object("Timeslot") %>%
  enter_object(time$key[i])%>%
    spread_values(Predicted_occupancy= jstring("Logistic_occupancy"), Expected_people= jnumber("People_estimate"), Max_people = jnumber("Max_people_estimate"), Min_people = jnumber("Min_people_estimate"))%>%
  select(Predicted_occupancy, Expected_people, Max_people, Min_people)
  
  if (i == 1) { 
              occupancy_table <- occupancy #create dataframe room occupancy
         } 
         else {
              occupancy_table <- rbindlist(list(occupancy_table, occupancy))#add new rows to the dataframe
         }
  
  }
  library(stringr)
  occupancy_table <-cbind(Time=str_sub(time$key, -8,-4), occupancy_table)
  occupancy_table <- occupancy_table[order(Time),] 
  occupancy_table$Predicted_occupancy <- as.factor(occupancy_table$Predicted_occupancy)
levels(occupancy_table$Predicted_occupancy)
# [1] "a" "b" "c" "d"

occupancy_table$Predicted_occupancy <- factor(occupancy_table$Predicted_occupancy, levels = c(" Low ", " Mid_Low ",  " Mid_High ",  " High " ))

  

```
**Day**: `r date_input`  
**Campus**:`r details$Campus`  
**Building**: `r details$Building`  
**Floor number**: `r details$Floor`  
**Room**: `r room`  
**Capacity of the room**: `r details$Capacity`   
**Plug-friendly**: `r details$Plug_friendly`  


This document summarises the occupancy of the room estimated using the predicted values gathered through our models for the day requested.  

The data are summarised by the following table and graphs.  

```{r pander, echo = FALSE, message=FALSE, warning=FALSE} 
library(pander)
panderOptions("digits", 4)
pander(occupancy_table)

```


```{r, echo=FALSE, message=FALSE, fig.align='center'}
occupancy_plot <- ggplot(data=occupancy_table, aes(x=factor(Time), y=Predicted_occupancy, fill=Predicted_occupancy)) + geom_bar(stat="identity")+ scale_fill_manual(values=c( "orangered","gold","darkcyan", "blue"))+ theme_bw()

count_plot <-ggplot(data=occupancy_table, aes(x=Time, y=Expected_people, group=1)) + geom_ribbon(data=occupancy_table,aes(ymin=Min_people, ymax=Max_people),alpha=0.2, fill="orangered3") + geom_line(color="orangered2")+ theme_bw() + geom_point(color="orangered2") 
multiplot(occupancy_plot, count_plot, cols=2)
```


