---
title: "Logistic_regression_report"
author: "Silvia Saloni"
date: "Thursday, July 28, 2016"
output:
  pdf_document:
    fig_height: 8
    fig_width: 14
---

#Introduction

The aim of the project was to be able to find a model that best predict the relationship between:
* the number of people counted with the survey for a given class at a particular hour 
* Wi-fi log counted in that room at that hour

This will allow to see whether Wi-Fi log is a good predictor for estimating occupancy in a classroom.

Given the fact that the target feature was a percentage we tried to run a logistic regression to predict when the room was more likely to be empty or occupied. 

Below we describes step by step all the analysis performed.

## ANALYSIS

### DATABASE CONNECTION AND DATASET

```{r, echo=FALSE, warning=FALSE}
#load all the libraries needed
library(RMySQL) #package for communicate with MySQL database
library(ggplot2) #package for making graphs
library(GGally)
library(nlme)
library(caret) # for splitting the database
library(DAAG)#for k-fold validation on linear and logistic
library(boot)#for k-fold validation on glm
library(plyr)
library(reshape2)
source("http://peterhaschke.com/Code/multiplot.R") 
```

First of all, we set up the connection to the database, using the following code:

```{r}
connection <- dbConnect(MySQL(),user="root", password="",dbname="who_there_db", host="localhost")
```  
Then we made a query to the database, in order to get all the groundth truth data collected in room B.002, B.004 and B.006 from 9 to 17 and the correspondent Wi-Fi Log measured in that time frame and rooms. 

```{r, echo=FALSE, warning=FALSE}
query <-"SELECT W.`Room_Room_id` as Room, W.`Date`, HOUR( W.Time ) as Time, T.`Module_Module_code` as Module, M.`Course_Level`,T.`Tutorial`, T.`Double_module`, T.`Class_went_ahead`, R.`Capacity`, G.`Percentage_room_full`, AVG(W.`Associated_client_counts`) as Wifi_Average_clients, MAX(W.`Authenticated_client_counts`) as Wifi_Max_clients FROM Room R, Wifi_log W, Ground_truth_data G, Time_table T, Module M WHERE W.Room_Room_id = R.Room_id AND G.Room_Room_id = W.Room_Room_id AND W.Date = G.Date AND HOUR( W.Time ) = HOUR( G.Time ) AND HOUR( W.Time ) = HOUR( T.Time_period ) AND T.Date = W.Date AND T.Room_Room_id = W.Room_Room_id AND M.`Module_code` = T.`Module_Module_code` GROUP BY W.Room_Room_id, HOUR( W.Time ) , W.Date"

#select the data based on the query and store them in a dataframe called Analysis table
AnalysisTable <-dbGetQuery(connection, query)  
```

The dataset created had in total 216 rows and it will allow us to explore if Wi-Fi log can estimate precisely if a room is empty or occupied during a certain hour.

As a **target features** for the logistic regression we decided to create a binary categorical features with 2 levels using the percentage of the room full. In particular, we consider as empty all the cases in which the occupancy of the room was 0 and occupied when the occupancy of the room was higher than 0. Empty was indicated as 0, while Occupied as one. 
```{r, echo=FALSE, warning=FALSE}
#create the new column for creating the categorical target feature
AnalysisTable$Binary_Occupancy[AnalysisTable$Percentage_room_full <= 0] <- "0"
AnalysisTable$Binary_Occupancy[AnalysisTable$Percentage_room_full>0] <- "1"
AnalysisTable$Binary_Occupancy <- factor(AnalysisTable$Binary_Occupancy)
```
As response variables or feature we considered Wi-Fi logs, which were summarised either as average of the logs counted for each room and for each hour or as maximum of the logs measured for each room and for each hour. 

Together with the Wi-Fi log, we included in the data set the following features: 

* **Date**, which we did not use in this analysis, because they just cover 2 weeks of Novemeber, but for future analyses they can be used to group observations by seasons or semesters or to finds seasonal trends for time series analyses.

* **Time**, which will be explored either as continous variable and as categorical to explore if the time of the day can have an affect on the Wi-Fi log. To do so we, bin the time in 4 ranges: early morning (9-11), late morning (11-13), early afternoon (13-15) and late afternoon (15-17). This will allow us to see if the Wi-Fi log accuracy was changing during the day. For example, it is more likely that all the electronic devices are fully powered early in the morning and consequently the Wi-fi log data can be more accurate or overestimating the occupancy of the room (i.e. more than one device per person). On the contrary in the afternoon, the devices may be more likely to be out of battery and it is possible that there are less devices in the room.
```{r, echo=FALSE, warning=FALSE}
AnalysisTable$Factor_Time <-cut(AnalysisTable$Time, breaks = 4, right=FALSE, labels=c('Early Morning','Late Morning','Early Afternoon','Late Afternoon' ))
```

* **Module**, which we are not going to include in the analysis because the majority of the module present are for computer science, but for future analyses it will be possible to explore if the Wi-fi log accuracy in predicting the occupancy change acroos the courses. Science course or computer science course will more likely to use electronic devices during lecture than art students.

* **Course level**, which can indicate us whether electronic devices will be less used during different course level. For example, first and second level courses can be less practical and therefore laptop are not needed and that can decrease the number of devices connected. On the other hand, undergraduate might be more distracted during lecture and look at their phones during lectures. This will result in an increase of connection in that hour.

```{r, echo=FALSE, warning=FALSE}
AnalysisTable$Course_Level <- factor(AnalysisTable$Course_Level)
```

* **Tutorial**, which can affect the number of logged people. First of all, because tutorial divided the room in 2 and therefore there will be measured less people than expected. 
```{r, echo=FALSE, warning=FALSE}
AnalysisTable$Double_module <- factor(AnalysisTable$Double_module)
AnalysisTable$Class_went_ahead <- factor(AnalysisTable$Class_went_ahead)
```

* **Double_module**, categorical variable indicating wether in the class there are more than one module, increasing the number of people expected in the room.
```{r, echo=FALSE}
AnalysisTable$Double_module <- factor(AnalysisTable$Double_module)
```

* **Double_module**, categorical variable indicating wether in the class went ahead to check for false positive.
```{r, echo=FALSE}
AnalysisTable$Class_went_ahead <- factor(AnalysisTable$Class_went_ahead)
```

The resulting data set is printed below:
```{r}
head(AnalysisTable)
```

## DATA QUALITY REPORT  

Before running any analyses, we carried out the data quality report to check for any issue related to the variable (e.g. outlier, skewed distribution, NaN values) and solutions we will implement to solve them.

Initially set all the categorical variables as factors and then we printed the descriptive statistic for all the features.
```{r}
summary(AnalysisTable)
```
From this we could see that NaN values were not present in the data set. 
We could notice that the observations for the features Tutorials and Double_model were not even distributed across the 2 levels of the variables. In fact, only 6 observations were present for tutorial class and for double module class. Therefore, we decided to discard both the features, because they will be not informative for the analysis. Similarly for the feature class_went_ahead the majority of the lectures did occur and we decided to discard it.
Furthermore for the variables Wifi_Average_clients and Wifi_Max_clients, it seems that there are few outliers, since the median is lower than the mean and the max values are far higher than the mean values. We will going to explore this issues with histogram and boxplots.

### Exploratory graphs

For exploring possible issues related with the continuous variables we plotted histograms and boxplots.

### Histograms

```{r, echo=FALSE,warning=FALSE}
#histogram for showing the count in each bin for the Maximum number of clients
histo1 <- ggplot(AnalysisTable, aes(x = Wifi_Max_clients)) + geom_histogram(binwidth = 10,  col="red", aes(fill=..count..)) + scale_fill_gradient("Count", low = "yellow", high = "red") +theme_bw()+theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 

#histogram for showing the count in each bin for the Average number of clients
histo2 <- ggplot(AnalysisTable, aes(x = Wifi_Average_clients)) + geom_histogram(binwidth = 10,  col="red", aes(fill=..count..)) + scale_fill_gradient("Count", low = "yellow", high = "red") +theme_bw()+theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 

#histogram for showing the count in each bin for each hour of the day
histo3 <- ggplot(AnalysisTable, aes(x = Time)) + geom_histogram(binwidth = 2,  col="red", aes(fill=..count..)) + scale_fill_gradient("Count", low = "yellow", high = "red") +theme_bw()+theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 

#plot all the histograms in one window
multiplot(histo1, histo2, histo3, cols=2)
```

<br><br><br>
Form the histograms we could see that the distribution of the feature Wifi Maximum_client (i.e. the Maximum number of devices logged in one hour lecture) was skewed to the left, indicating that the in the majority of the lecture were counted no more than 40 people. Furthermore, we could see that there are potential outliers (values > 150). Similar pattern was observed for the feature Wifi_Average_clients.
Feature times had as well a skewed distribution, suggesting that the majority of the lectures were concentrating during the early morning and they were decreasing towards the afternoon.
<br><br><br>
#### Boxplots
<br>
  
```{r, echo=FALSE, warning=FALSE}
#make the boxplot for continuous variable
#box plot for the counted client varable
box1 <- ggplot(AnalysisTable, aes(x = factor(0), y = Wifi_Average_clients)) + geom_boxplot() + xlab("Average counted clients") + ylab("")+ scale_x_discrete(breaks = NULL)  + theme_bw()+theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 

#box plot for the maximum number of clients variable
box2 <- ggplot(AnalysisTable, aes(x = factor(0), y =Wifi_Max_clients)) + geom_boxplot() + xlab("Maximum counted clients") + ylab("")+ scale_x_discrete(breaks = NULL) + theme_bw()+theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 

#box plot for the Time continuous variable
box3 <- ggplot(AnalysisTable, aes(x = factor(0), y = Time)) + geom_boxplot() + xlab("Time") + ylab("")+ scale_x_discrete(breaks = NULL) + theme_bw()+theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 
#plot all the boxplots in one window
multiplot(box1, box2, box3, cols=2)
```  
<br><br>

From the boxplots, all the trends observed in the histograms were confirmed.

For categorical variables we plot bar plot graphs.  

#### Bar plots  

```{r, echo=FALSE, warning=FALSE}

bar1 <- ggplot(AnalysisTable, aes(x = Room)) + geom_bar(fill="orangered2")+ theme_bw()+theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 

#bar plot for the categorical variable: Course level
bar2 <- ggplot(AnalysisTable, aes(x = Course_Level)) + geom_bar(fill="orangered2")+ theme_bw()+theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 

#bar plot for the categorical variable: Time as factor
bar3 <- ggplot(AnalysisTable, aes(x = Factor_Time)) + geom_bar(fill="orangered2")+ theme_bw()+theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 

#plot all the barplots in one window
multiplot(bar1, bar2, bar3, cols=2)
```
<br><br>
From the barplots, we could see that observations were equally distributed across all the levels of the feature Room and Factor Time. On the contrary, there were more observations for non lectures and level 3 courses and there were more occupied classes. No issues were detected for those features.  

#### Summary
<br><br>

Features               | Issues                                   | Planned Solution             
-----------------------|------------------------------------------|----------------------------
Room                   |None                                      |None
Time                   |Distibution skewed to the left            |To solve during analysis 
Factor Time            |None                                      |None
Course level           |None                                      |None
Tutorial               |Uneven representation of the level        |Discarded from the analysis
Double Module          |Uneven representation of the level        |Discarded from the analysis
Class went ahead       |Uneven representation of the level        |Discarded from the analysis
Wifi Average clients   |Distibution skewed to the left \& outliers|To solve during analysis
Wifi Maximum clients   |Distibution skewed to the left \& outliers|To solve during analysis
Binary Occupancy       |None                                      |To solve during analysis

<br><br>

## FEATURES AFFECTING THE TARGET FEATURE  

The next step of the analysis was to see which feature really affect the target feature for deciding which features we would include into the model.  
 
For the continuous features we used box plots, while for categorical we use bar plots.

```{r, echo=FALSE, warning=FALSE}
pairbox1 <- ggplot(AnalysisTable, aes(x = Binary_Occupancy, y =Wifi_Average_clients)) + geom_boxplot()+ theme_bw()+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 

pairbox2 <- ggplot(AnalysisTable, aes(x = Binary_Occupancy, y = Wifi_Max_clients)) + geom_boxplot() + theme_bw()+theme( panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 

pairbox3 <- ggplot(AnalysisTable, aes(x =Binary_Occupancy, y = Time )) + geom_boxplot() + theme_bw()+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 

multiplot(pairbox1, pairbox2, pairbox3, cols=3)
```
<br><br><br>

From the boxplots we could see that when the room was indicated as empty either the average and the maximum Wi-Fi logs were close to zero. On the contrary, when the room was occupied the average and the maximum Wi-Fi logs were different from zero. The difference between the 2 levels will probably be higher without the outliers. Therefore, we can conclude that either average Binary_Occupancy and maximum wifi counted clients are a good predictor of the binary occupancy and we are going to run 2 models: one for exploring the relationship between Binary_Occupancy and Average counted clients and another for Survey counted clients and Maximum counted clients. 

## Barplots 

```{r, echo=FALSE,warning=FALSE}
barpair1 <-ggplot(AnalysisTable, aes(x = Room, fill = Binary_Occupancy)) + geom_bar(position = "dodge")+ scale_fill_manual(values=c( "cyan4","orange"))+theme_bw()+theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

barpair2 <-ggplot(AnalysisTable, aes(x = Factor_Time, fill =Binary_Occupancy)) + geom_bar(position = "dodge")+ scale_fill_manual(values=c( "cyan4","orange"))+theme_bw()+theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

barpair3 <-ggplot(AnalysisTable, aes(x = Course_Level, fill = Binary_Occupancy)) + geom_bar(position = "dodge")+ scale_fill_manual(values=c( "cyan4","orange"))+theme_bw()+theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))


multiplot(barpair1, barpair2, barpair3, cols=2)
```

From the barplots, we could see that the occupancy was not affected by the time of room. On the contrary, the occupancy of the room increase for level3 and 4 courses, while was lower for level1, 2 and 5, which in our case represents career talk or CS meeting.
The occupancy of the room in late monrning and late afternoon seemed to be higher.
Therefore, in our model we are going to explore whether course level and time of the day affect the chance of a room to be empty or occupied.

### Interacting effect on the target feature  
The last step before the regression was to explore the interacting effect of the features on the target features. 

First of all we explore the interactive effect between Average conted clients with all the other features using bar plots.

```{r, echo=FALSE,warning=FALSE}
#Graph exploring interactive effect between Wifi_Average_clients and Time on binary occupancy 
pair1 <- ggplot(AnalysisTable, aes(x = Factor_Time, y =Wifi_Average_clients, fill = factor(Binary_Occupancy))) + geom_bar(position = "dodge", stat="identity")+ scale_fill_manual(values=c( "darkblue","cyan4", "yellow"))+theme_bw()+theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

#Graph exploring interactive effect between Wifi_Average_clients and Course level on binary occupancy 
pair2 <- ggplot(AnalysisTable, aes(x = Course_Level, y =Wifi_Average_clients, fill = factor(Binary_Occupancy))) + geom_bar(position = "dodge", stat="identity")+ scale_fill_manual(values=c( "darkblue","cyan4", "yellow", "orange", "blue", "red"))+theme_bw()+theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
#Graph exploring interactive effect between Wifi_Max_clients and Time on binary occupancy
pair3 <- ggplot(AnalysisTable, aes(x = Factor_Time, y =Wifi_Max_clients, fill = factor(Binary_Occupancy))) + geom_bar(position = "dodge", stat="identity")+ scale_fill_manual(values=c( "darkblue","cyan4", "yellow"))+theme_bw()+theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

#Graph exploring interactive effect between Wifi_Max_clients and Course level on binary occupancy
pair4 <- ggplot(AnalysisTable, aes(x = Course_Level, y =Wifi_Max_clients, fill = factor(Binary_Occupancy))) + geom_bar(position = "dodge", stat="identity")+ scale_fill_manual(values=c( "darkblue","cyan4", "yellow", "orange", "blue", "red"))+theme_bw()+theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

multiplot(pair1, pair2, pair3, pair4, cols=2)
```

From the graphs we could see that the average and the maximum number of devices counted with the Wifi-logs were lower in rooms indicated as empty by the survey across all the level of the time factor, indicating that there was not an interaction between Wifi average count and time and between Wifi maximum count and time.
Similar pattern was found across all the levels of course level, suggesting that there was not interaction between Wifi average count and course levels and between Wifi maximum count and course levels.

For exploring the interaction between Time and course level we did the following bar plot:

```{r, echo=FALSE, warning=FALSE}
#Graph exploring interactive effect between time and course level on binary occupancy 
ggplot(AnalysisTable, aes(x=Factor_Time, fill = factor(Course_Level)) ) + facet_grid(Binary_Occupancy ~ .) + geom_bar(position = "dodge")+ scale_fill_manual(values=c( "darkblue","cyan4", "yellow", "orange", "blue", "red"))+theme_bw()+theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) 
```
From this graph we cannot detect any clear pattern, therefore, we are not going to explore the interactive effect between Time and course level.

Consequently we are going to explore the following 2 models:
* $Binary_Occupancy ~ Average Wifi occupancy + Time + Course_Level$
* $Binary_Occupancy ~ Maximum Wifi occupancy + Time + Course_Level$
<br>

# Analysis
<br>
For running the logistic regression we performed k fold cross validation with the aim of testing if the selected model can be applied to new datasets. K fold validation was preferred over the validationset approach and the Leave Out Cross Validation (LOOCV), because it is more robust and more accurate in estimating the test error. The Validation set approach tends to give an over estimate of the test error and the test error is dependent on the observations included randomly in the test set. Furthermore, the LOOCV tends to provide a test error with a high variance, because the folds used to calculating it are correlated among each other. 
In particular in this analysis we are going to perform a 10 fold cross validation, which is pretty standard.
<br>

# Results
The model $Binary_Occupancy ~ Maximum Wifi occupancy + Time + Course_Level$ was very accurate, as showed below:
```{r, warning=FALSE}
Logitmodel1 <- glm(Binary_Occupancy ~Wifi_Max_clients  + Factor_Time + Course_Level,family=binomial,data =AnalysisTable)

CVbinary (Logitmodel1, nfolds= 10)
```

Also the second model, $Binary_Occupancy ~ Average Wifi occupancy + Time + Course_Level$ had a good accuracy, but lower than the previous one. 

```{r,  warning=FALSE}
Logitmodel2 <- glm(Binary_Occupancy ~Wifi_Average_clients + Factor_Time + Course_Level,family=binomial,data =AnalysisTable)

CVbinary (Logitmodel2, nfolds= 10)
```

Therefore, we selected as best logistic model the first one, which suggested that Max Wifi occupancy, Time and Course_Level were good predictor of the occupancy of the room.
